BDP draft paper

### 2016-08-03 Rebeccas_location_Code2/StopData_Maps.Rmd
draft paper

# Get Shapefiles
```{r}
locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_blockgroups_2010", verbose = TRUE)
summary(blocks)
print(proj4string(blocks))
blocks.polygons <- blocks@polygons
blocks.data <- blocks@data
blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

@@

# Get Shapefiles Again
```{r}
locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
summary(blocks)
print(proj4string(blocks))
# blocks.polygons <- blocks@polygons
# blocks.data <- blocks@data
# blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

@@

## Population by census block group:
```{r}
shpfile <- "Census_tracts_2010.shp"
sh <- readShapePoly("Census_Tract_Polygons2010/Census_tracts_2010.shp")
plot(sh)
sh@data$ID <- as.numeric(sh@data$ID)

@@

# Scale the total population to the palette
pop <- sh2@data$TotalPop.x
cols <- (pop - min(pop))/diff(range(pop))*127+1
plot(sh, col=cols)
#points( mergedf$long, mergedf$lat, col="black", cex=.6 )

>>>>

# Get Shapefiles
```{r}
#locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Block_Group_Polygons2010"
blocks <- readOGR("Census_Block_Group_Polygons2010","Census_blockgroups_2010", verbose = TRUE)

blocks.polygons <- blocks@polygons
blocks.data <- blocks@data
blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

@@

# Get Shapefiles Again
```{r}
locationCensusFiles <- "C:/Users/Rebecca/Dropbox/School/2016_su_School/stat133/finalproject_angry_ladies/Rebeccas_location_Code2/Census_Tract_Polygons2010"

locationCensusFiles <- "Census_Tract_Polygons2010"
blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)

locationCensusFiles <- "Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
summary(blocks)
print(proj4string(blocks))
# blocks.polygons <- blocks@polygons
# blocks.data <- blocks@data
# blocks.plot.order <- blocks@plotOrder
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)
```

@@

## Population by census block group:
```{r}
shpfile <- "Census_tracts_2010.shp"
sh <- readShapePoly("Census_Tract_Polygons2010/Census_tracts_2010.shp")
#plot(sh)
sh@data$ID <- as.numeric(sh@data$ID)

@@

# Scale the total population to the palette
pop <- sh2@data$TotalPop.x
#str(sh2)
#sh2[1]
#sh2@data$TotalPop.x
#pop <- c(2925,3144,4196,4658,3558,3387,2007,3198,3563,1992,1756,3632,3118,3844,3716,3647,3640,8368,3633,5659,2642,2794,1593,2685,1552,2020,2172,3964,4396,4885,1215,4336,4685)

cols <- (pop - min(pop))/diff(range(pop))*127+1
plot(sh, col=cols)
#points( mergedf$long, mergedf$lat, col="black", cex=.6 )

### 2016-08-03 final_paper/draft.Rmd
draft paper

---
title: "BPD_draftpaper"
author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
date: "2 August 2016"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(readr)
library(plyr)
library(dplyr)
library(ggvis)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
```

@@

* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)

@@

	For the stop data, there were 16,255 incidents assessed by the Berkeley police. In each case, the call date and time, location, incident type, and disposition(s) were recorded. The cases typically had a six character disposition, with each character conveying race, gender, age, reason, enforcement, and car search,  respectively, for each subject involved in the incident, although there were additional dispositions, ranging from one to three characters,  that could be input and conveyed other information. In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns. To improve readability, the acronyms and abbreviations used in the data set were mutated to convey their meanings  (will do soon).

@@

```{r, echo=FALSE}
# Read data
stop_by <- read_csv("cleaned_stop_data.csv")

@@

stop_by <- stop_by %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))
```

### About the Berkeley PD Stop Data

The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department. Raw data set is downloaded from City of Berkeley Open Data website and provided by Berkeley Police Department.

Analysis in this part focuses on the information provided by the dispositions variable. The dispositions variable can be divided into three types. The first type only includes more than one six-character person message. Each character in each person message represents the race, gender and age range of the person, reason and enforcement of the stop, and the car search information during the stop. The second type of the dispositions includes only additional dispositions, such as P for primary cases. The third type of dispositions includes both person information and additional dispositions.

For the analysis in this part, only person information for all the people in each incident and the Call Time/Date of each incident are analyzed.

@@

```{r, echo=FALSE, eval=FALSE}
stop_race_total_table <- stop_by %>%
    group_by(Race) %>%
    summarise(totaLCount = n())
stop_race_agerange_table <- stop_by %>%
    group_by(Race, AgeRange) %>%
    summarise(count = n())
stop_race_table <- left_join(stop_race_agerange_table, stop_race_total_table, by = "Race")
stop_race_table <- stop_race_table %>%
    mutate(percentage = count/totaLCount) %>%
    select(-count)
stop_race_table <- stop_race_table %>%
    spread(key = AgeRange, value = percentage)
stop_race_table
```

@@

Race  | Count | 0-18 | 18-29 | 30-39 | 40+
----  |------ |------|-------|-------|-------
A     | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
B     | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
H     | 1676  | 3.22%| 46.12%| 29.18%| 21.47%
O     | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
W     | 5454  | 1.72%| 29.45%| 25.39%| 43.44%

#### Picture: count of person recorded of each race in each hour
```{r, echo=FALSE}
stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = ..count..)) +
  labs(y = "Count") +
  facet_wrap(~Race)
stop_time_percentage
```

(1) Among the 14291 personnel information recorded, white people contributes the largest percentage. 38.16% of all records, namely 5454 people are white. black people contributes the second largest percentage. 32.44% of all records, namely 4636 people are black.

(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race, which makes sense in daily life.

(3) Among white people and black people recorded by the BPD, people aged greater than 40  contribute the largest percentage, which are 43.44% and 37.66% respectively.

@@

(5) Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.

@@

```{r, eval = FALSE, echo=FALSE}
stop_hour <- stop_by %>%
  group_by(Hour) %>%
  summarise(count = n())
stop_hour
```

@@

#### Table: Probability of arrested by BPD of a specific age range for a specific reason
P     | I     | K     | R     | T     | W     
------|-------|-------|-------|-------|-------
0-18  | 17.95%| 67.57%| 13.95%| 3.24% | 50.00%  
18-29 | 8.45% | 12.50%| 3.87% | 1.82% | 32.35%  
30-39 | 6.32% | 25.00%| 5.04% | 1.49% | 26.92%  
40+   | 7.87% | 23.86%| 6.13% | 1.41% | 31.82%

@@

```{r, echo=FALSE}
## Read in the Data:
#finaldf <- readRDS( 'StopData_finaldf.rds' )
mergedf <- readRDS( 'StopData_merged.rds' )
```

```{r, echo=FALSE}
## Location info:
latmax <- max( mergedf$lat, na.rm = TRUE ) 
latmin <- min( mergedf$lat, na.rm = TRUE )
lonmax <- max( mergedf$long, na.rm = TRUE ) 
lonmin <- min( mergedf$long, na.rm = TRUE )
latvals <- c( latmin, latmax )
lonvals <- c( lonmin, lonmax )
```

```{r, echo=FALSE}
## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)

## MAP 1:All BPD Stops Density, 2015-2016
```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
# a contour plot
ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = mergedf, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme (panel.grid.major = element_blank (), # remove major grid
		       panel.grid.minor = element_blank (),  # remove minor grid
		       axis.text = element_blank (), 
		       axis.title = element_blank (),
	         axis.ticks = element_blank ()) + 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)

## MAP 2: BPD Stop Contour Map of Berkeley by Race
```{r, fig.height = 5, fig.width = 10, echo=FALSE}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 100) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

## MAP 3: BPD Stop Contour Map of Berkeley by Age Range
```{r, fig.height = 8, fig.width = 10, echo=FALSE}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "Stop Density",
    low = "white", mid = "orange", high = "red", midpoint = 25) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

We found that the

>>>>

---
title: "BDP draft paper"
author: "Amy Zhu; Mengyu Li; Rebecca Reus; Shangjun Jiang"
date: "2 August 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sp)
library(ggmap)
library(tidyr)
library(readr)
library(plyr)
library(dplyr)
library(ggvis)
library(tidyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(readr)
library(data.table)
library(rworldmap)
library(maps)
library(mapdata)
library(maptools)
library(scales)
library(RgoogleMaps)
library(tmap)
library(sp)
library(rgdal)
library(rgeos)
library(RColorBrewer)
library(grid)
library(gridExtra)
```

@@

* [Stop Data (16,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Stop-Data/6e9j-pj9p)
* [Arrest (200)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Arrests/xi7q-nji6)
* [Jail Bookings (250)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Log-Jail-Bookings/7ykt-c32j)
* [Calls for Service (4,000)](https://data.cityofberkeley.info/Public-Safety/Berkeley-PD-Calls-for-Service/k2nh-s5h5)
* [Berkeley Census Data](https://data.cityofberkeley.info/)

@@

	For the stop data, there were 16,255 incidents assessed by the Berkeley police. In each case, the call date and time, location, incident type, and disposition(s) were recorded. The cases typically had a six character disposition, with each character conveying race, gender, age, reason, enforcement, and car search,  respectively, for each subject involved in the incident, although there were additional dispositions, ranging from one to three characters,  that could be input and conveyed other information. In order to prepare the data for exploration, the cleaning process included changing the call date and time to the lubridate format, transforming the addresses into longitude and latitude referencing Google Maps, and separating the information on dispositions into separate row entries for each individual assessed in a case and splitting it further by isolating “other” and six character dispositions into different columns.

@@

```{r, echo=FALSE}
# Read data

############################## stop data ############################## 
stop_by <- read_csv("cleaned_stop_data.csv")

@@

stop_by <- stop_by %>%
  mutate(Hour = as.integer(hour(Call.Date.Time))) %>%
  mutate(Day = as.factor(as.integer(wday(Call.Date.Time))))

############################## location data ############################## 
mergedf <- readRDS( 'StopData_merged.rds' )

## Location info:
latmax <- max(mergedf$lat, na.rm = TRUE) 
latmin <- min(mergedf$lat, na.rm = TRUE)
lonmax <- max(mergedf$long, na.rm = TRUE) 
lonmin <- min(mergedf$long, na.rm = TRUE)
latvals <- c(latmin, latmax)
lonvals <- c(lonmin, lonmax)

## Make the variables factors:
mergedf$CarSearch <- factor(mergedf$CarSearch)
levels(mergedf$CarSearch) <- c("No Search", "Search")

### BerkeleyMap 
berkMap = map = get_map(location = c( lon = mean(lonvals), lat = mean(latvals) ), zoom = 14)

############################## census data ##############################

blocks <- readOGR("Census_Tract_Polygons2010","Census_tracts_2010", verbose = TRUE)

locationCensusFiles <- "Census_Tract_Polygons2010"
blocks <- readOGR(locationCensusFiles,"Census_tracts_2010", verbose = TRUE)
b2 <- spTransform(blocks, CRS("+proj=longlat +datum=WGS84"))
b3 <- fortify(b2)

############################## arrest data ############################## 
arrest <- read.csv("Berkeley_PD_Log_-_Arrests.csv")

############################## jail data ############################## 
jail <- read.csv("Berkeley_PD_Log_-_Jail_Bookings.csv")

############################## multiplot function ############################## 
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

We found the census data from the Berkeley open data website and created a headmap of the Berkeley population. Similarly, we created a heat map based on the stop data.  Interestingly, we found that the most dense place (around north Berkeley) is a relatively safe place. The place where people are more likely to be stopped (downtown berkeley area) is less dense. Since Downtown Berkeley area is a transportation hub, numerous people come and go around this area. Though there are more people living north Berkeley, mainly residents will visit the area. Therefore, the most likely place to get stopped is not the most dense living place.

### Heat Map 1: Berkeley Census data.
```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = b3, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme (panel.grid.major = element_blank (), # remove major grid
		       panel.grid.minor = element_blank ()  # remove minor grid
		       ) + 
  ggtitle("Population Density") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)
```

### Heat Map 2: All BPD Stops Density, 2015-2016
```{r, fig.width=10, fig.asp=0.8, echo=FALSE}
# a contour plot
ggmap(berkMap) +
  stat_density2d(aes(x = long, y = lat, fill= ..level.., alpha = .2* ..level..),
    size = 2, bins = 5, data = mergedf, geom = "polygon") +
  scale_fill_gradient(low = "black", high = "red") +
    theme (panel.grid.major = element_blank (), # remove major grid
		       panel.grid.minor = element_blank ()  # remove minor grid
		       )+ 
  ggtitle("All BPD Stops Density, 2015-2016") +
  labs(alpha = element_blank())+
  guides(alpha = FALSE)

Then we explored the stop data by age range (0-18; 18-29; 30-39; and 40+) and data. Similar to the all stop data density, the area that is most likely to be stopped is the same: Downtown Berkeley. The differences in age range and race don't play a crucial role in the possibility of being stopped in Berkeley. These two analysis confirmed our explanations for the all stop data heat map.

### Heat Map 3: BPD Stop Contour Map of Berkeley by Age Range
```{r, fig.height = 8, fig.width = 10, echo=FALSE}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "Stop Density",
    low = "white", mid = "orange", high = "red", midpoint = 25) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap(~ AgeRange) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Age Range") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

Note: 1 refers to age 0-18; 2 refers to age 18-29; 3 refers to age 30-39; and 4 refers to age 40+.

## Heat Map 4: BPD Stop Contour Map of Berkeley by Race
```{r, fig.height = 5, fig.width = 10, echo=FALSE}
ggmap(berkMap) +
  stat_density2d( aes(x = long, y = lat, fill = ..level.., alpha = ..level..),
    bins = I(5), geom = "polygon", data = mergedf ) +
  scale_fill_gradient2( "StopDensity",
    low = "white", mid = "orange", high = "red", midpoint = 100) +
  labs(x = "Longitude", y = "Latitude") + facet_wrap( ~ Race ) +
  scale_alpha(range = c(.2, .55), guide = FALSE) +
  ggtitle("BPD Stop Contour Map of Berkeley by Race") +
  guides(fill = guide_colorbar(barwidth = 1.5, barheight = 10))
```

The Berkeley PD stop data reflects accidents that have been visited by the Berkeley Police Department. After the initial exploration of the Berkeley stop data, we focus on on the information provided by the dispositions variable. The dispositions variable provides information on race, gender and age range, stop reason and enforcement of the stop, the car search information during the stop, and additional dispositions.

@@

@@

   Race     | Count | 0-18 | 18-29 | 30-39 | 40+
----------  |------ |------|-------|-------|-------
A (Asian)   | 1141  | 8.50%| 43.47%| 21.03%| 26.99%
B (Black)   | 4636  | 2.29%| 34.97%| 25.09%| 37.66%
H (Hispanic)| 1676  | 3.22%| 46.12%| 29.18%| 21.47%
O (Other)   | 1384  | 2.67%| 43.28%| 29.17%| 26.87%
W (White)   | 5454  | 1.72%| 29.45%| 25.39%| 43.44%

(1) Among the 14291 personnel information recorded, white people is the number one population group of being stopped by Berkeley Police. 38.16% of all records, namely 5454 people are white. black people contributes the second largest percentage. 32.44% of all records, namely 4636 people are black.

(2) The percentage of people aged from 0 to 18 of each race is less than 9%. People aged from 0 to 18 contributes the smallest percentage in each race. These teengers are looked after by their parents most of the time.

(3) Among white people and black people recorded by the BPD, people who are over 40 are the number one group that are being stopped, which are 43.44% and 37.66% respectively.

@@

Then we further analyzed the stop data with and without car searches by race. Since white people are the largest group to be stopped, we suspect that their possibility of car search will also be the highest. However, black people seem to be more likely to be car searched.

### MAP: BPD All Stops w/ Car Searches by Race 2015-2016
```{r, echo=FALSE}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, size = CarSearch, color = Race), data = mergedf, alpha = 0.2) + 
  scale_size_discrete( "Car Search", labels = c("No Search", "Search") ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD All Stops w/ Car Searches by Race 2015-2016") +
  facet_wrap(~ Race, ncol = 3) 
```

Even when we look at the arrest data, it is black people who are more likely to get arrest after being stopped.

```{r, echo=FALSE}
df <- subset(mergedf, as.character(Enforcement) == "Arrest")

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race), data = df, alpha = 0.5, size = 5) + 
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Stops to Arrests, by Race 2015-2016") +
  facet_wrap(~ Race) 
```

### Analysis of stop reasons 
#### Map: All stop data by reasons.
```{r, fig.height = 5, fig.width = 10, echo=FALSE}
ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Reason), data = mergedf, alpha = 0.7, size = 3) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + ggtitle("All BPD Stops, by Reason 2015-2016") +
    facet_wrap( ~ Reason , ncol = 2)
```

From the analysis of the stop reasons, we found that the number one reason that people are stopped is traffic reason. Then we explored traffic reasons within race. As shown, black people are the highest to be stopped and arrested.

#### Map: stop by only traffic reasons
```{r, fig.height = 5, fig.width = 10, echo=FALSE}
df <- subset(mergedf, as.character(Reason) == "Traffic")

ggmap(berkMap) +
  geom_point(aes(x = long, y = lat, colour = Race, size= Arrest), data = df, alpha = 0.5) + 
  scale_size_discrete( range = c(5, 2) ) +
  theme ( 
		panel.grid.major = element_blank (), # remove major grid
		panel.grid.minor = element_blank (),  # remove minor grid
		axis.text = element_blank (), 
		axis.title = element_blank (),
		axis.ticks = element_blank ()
		) + 
  ggtitle("BPD Traffic Stops Only, by Race 2015-2016") +
  facet_grid(~ Race)
```

Such contrast of being stopped, being car searched, and being arrested is further demonstrated by the arrest and jail data.

```{r, echo=FALSE}
p1 <- ggplot(na.omit(arrest))+
  geom_bar(aes(x=Age, 
               fill = Race))+
  facet_wrap(~Race)+
 labs(x= "Age",
       y = "Number of people",
       title = "Arrest Data")

p2 <- ggplot(na.omit(jail))+
  geom_bar(aes(x=Age,
               fill = Race))+
  facet_wrap(~Race)

multiplot(p1, p2,cols=2)

As we can clearly see from both graphs, black people are the largest group to be arrested and put into jail.

Then we looked at the counts of people recorded of different races in every hour.Black people contribute a especially high percentage of the incidents at night, while the white people contribute a especially high percentage of the incidents at the noon.

```{r, echo=FALSE}
stop_time_percentage <- ggplot(stop_by, aes(x = Hour, fill = AgeRange)) + 
  geom_bar(aes(y = ..count..)) +
  labs(y = "Count") +
  facet_wrap(~Race)+
  labs(title = "Counts of people recorded of each race in every hour")
stop_time_percentage
```

Such analysis made us further digged into the date information.

### Analysis of Day
```{r, echo=FALSE, eval = FALSE}
stop_day <- stop_by %>%
  ggvis(x = ~Day, fill = ~Race) %>%
  layer_bars() %>%
  add_tooltip(function(stop_by) stop_by$Race)+
  labs(title = "Counts of people recorded in every day in a week")
stop_day
```

####Picture: counts of people recorded in every day in a week
Day   | Count | A (Asian)| B (Black)| H (Hispanic) | O (Other)| W (White)
----  |------ |----------|----------|--------------|----------|------
1     | 1692  | 116      | 653      | 221          | 159      | 543
2     | 1665  | 149      | 543      | 182          | 166      | 625
3     | 2344  | 189      | 645      | 231          | 236      | 1043
4     | 2287  | 206      | 616      | 251          | 244      | 970
5     | 1941  | 152      | 592      | 223          | 168      | 806
6     | 2231  | 186      | 764      | 268          | 203      | 810
7     | 2131  | 143      | 823      | 300          | 208      | 657

(1) Among all the day in a week, the number of incidents happened on Sunday and Monday are much less than those happened through Tuesday and Saturday.

(2) The count of records of Asian people is much less than that of any other race of people in each day and the whole week.

(3) Asian people are less liable to commit an incident on Sunday, and more liable to commit an incident on Wednesday.

(4) Black and Hispanic people are less liable to commit an incident on Monday, and more liable to commit an incident on Saturday.

(5) White people are less liable to commit an incident on Sunday, and more liable to commit an incident on Tuesday and Wednesday, which is similar to Asian people.

@@

@@

#### Table: Probability of arrested by BPD of a specific age range for a specific reason
P     | I (Investigation) |K (Probation/ Parole)| R (Reasonable Suspicion)|T (Traffic)| W (Wanted) 
----  |------------------ |---------------------|-------------------------|-----------|-------
0-18  | 17.95%            | 67.57%              | 13.95%                  | 3.24%     | 50.00%  
18-29 | 8.45%             | 12.50%              | 3.87%                   | 1.82%     | 32.35%  
30-39 | 6.32%             | 25.00%              | 5.04%                   | 1.49%     | 26.92%  
40+   | 7.87%             | 23.86%              | 6.13%                   | 1.41%     | 31.82%

@@

## Conclusion

1. From analyzing the relationship between population density and stop data density, we found that he place that are the most likely to be stopped is not the place where most people live, but the transportation hub.
2. From various analysis of the data by race, such as the stop time by race, car search by race, we found that black people are the number one to be arrested and even put into jail. Surprisingly, they are not the number one to be stopped. White people are the largest group to be stopped. 
3. During daytime, young people and white people are more likely to be stopped. During nightime, people over 40 and black people are more likely to be stopped by police. 
4. Asian and female are more likely to be arrested by the BDP for probation/parole.
